The Companion Mass Distribution of extremely low-mass white dwarfs

### 2014-10-15 documents/letter/ms.tex
Added figshare description

% packages
\usepackage{amssymb,amsmath,amsbsy}
\usepackage{booktabs}
\usepackage{multirow}

@@

\begin{abstract}
Measuring the masses of companions to single-line spectroscopic binary stars
is (in general) not possible because of the unknown orbital plane inclination angles. 
Even when the mass of the visible star can be measured, only a lower limit can be placed
on the mass of the unseen companion. However, since the orbital geometry of the systems should be isotropically
distributed, for a large enough, unbiased sample, the 
companion mass distribution can be deconvolved from the distributions of observables.
In this work, we construct a hierarchical probabilistic model to infer properties of unseen 
companion stars given orbital period and projected radial velocity measurements of the primary. We use a mixture of 
two Gaussians to model the true companion mass distribution to represent 
populations of white dwarfs (WD) and neutron stars (NS). We run this model on mock data sets in order to test the
robustness of our inferences. We then apply our model to 55 of the 61 white dwarfs in 
the extremely low mass (ELM) WD survey. We find maximum a posteriori (MAP) parameters for
the WD companion population of $\mu_{\rm WD} = 0.73~\Msun$ and $\sigma_{\rm WD} = 0.26~\Msun$ 
with a NS fraction consistent with zero, $f_{\rm NS} = 0\%$. However, the marginal posterior 
distribution over NS fraction has a significant tail up to $f_{\rm NS} \approx 12\%$.
We make samples from the posterior distribution publicly available and provide a  
script for computing the probability that a ... in the context of our model.
\end{abstract}

@@

\section{Test Cases} \label{sec:tests}
To test the performance of this Gaussian mixture model, we apply it to three separate sets of mock data (described in detail below), each with 100 ``observed'' systems. These data are generated by randomly drawing the companion mass, $M_2$, from a particular distribution, then using a randomly drawn inclination angle and primary mass, $M_1$, to compute the mass function, $\mf$. We draw primary masses from a uniform distribution, $M_1 \sim \mathcal{U}(0.2,0.4)~\Msun$. We apply the same Gaussian mixture model to all test datasets to infer the parameters of the white dwarf mixture component and neutron star fraction.

@@

\section{The ELM Sample}

@@

The probability of any particular ELM WD having a NS companion will vary depending on the NS fraction of the sample. This is not taken into account in \eqn\ref{eq:P_NS_approx}. {\bf TEXT ON USING PYTHON SCRIPT FOR DETERMINING P-NS FOR FUTURE DISCOVERED LMWDs.}

>>>>

% packages
\usepackage{amssymb,amsmath,amsbsy}
\usepackage{booktabs}
\usepackage{multirow}
\usepackage{url}

@@

\begin{abstract}
Measuring the masses of companions to single-line spectroscopic binary stars
is (in general) not possible because of the unknown orbital plane inclination angles. 
Even when the mass of the visible star can be measured, only a lower limit can be placed
on the mass of the unseen companion. However, since the orbital geometry of the systems should be isotropically
distributed, for a large enough, unbiased sample, the 
companion mass distribution can be deconvolved from the distributions of observables.
In this work, we construct a hierarchical probabilistic model to infer properties of unseen 
companion stars given orbital period and projected radial velocity measurements of the primary. We use a mixture of 
two Gaussians to model the true companion mass distribution to represent 
populations of white dwarfs (WD) and neutron stars (NS). We run this model on mock data sets in order to test the
robustness of our inferences. We then apply our model to 55 of the 61 white dwarfs in 
the extremely low mass (ELM) WD survey. We find maximum a posteriori (MAP) parameters for
the WD companion population of $\mu_{\rm WD} = 0.73~\Msun$ and $\sigma_{\rm WD} = 0.26~\Msun$ 
with a NS fraction consistent with zero, $f_{\rm NS} = 0\%$. However, the marginal posterior 
distribution over NS fraction has a significant tail up to $f_{\rm NS} \approx 12\%$.
We make samples from the posterior distribution publicly available and provide a  
script for computing the companion mass distribution and the NS probability for newly discovered low mass WDs given a measured WD mass and mass function.
\end{abstract}

@@

\section{Applying Our Model} \label{sec:tests}
To test the performance of this Gaussian mixture model, we first apply it to three separate sets of mock data (described in detail below), each with 100 ``observed'' systems. These data are generated by randomly drawing the companion mass, $M_2$, from a particular distribution, then using a randomly drawn inclination angle and primary mass, $M_1$, to compute the mass function, $\mf$. We draw primary masses from a uniform distribution, $M_1 \sim \mathcal{U}(0.2,0.4)~\Msun$. We apply the same Gaussian mixture model to all test datasets to infer the parameters of the white dwarf mixture component and neutron star fraction.

@@

\subsection{The ELM Sample}

@@

To determine the probability of any particular LMWD of hosting a NS, we make our model posteriors publicly available on fig{\bf share} \footnote{\url{http://figshare.com/authors/Jeff\_Andrews/644768}}. We further provide a {\tt python} script that calculates $P_{\rm NS}$ and the mass distribution for a WD companion for any LMWD with a measured $M_1$ and $\mf$. This script can be applied to newly discovered LMWDs as well as those already in the ELM sample.
